---
title: "Assigment - kNN DIY"
author:
  - Ryan Kokke - Author
  - Audrya Kerenhappukh - Reviewer         
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
   html_notebook:
    toc: true
    toc_depth: 2
---


```{r}
library(tidyverse)
library(googlesheets4)
library(class)
library(caret)
```

---

For this assignment a dataset containing laboratory values of blood donors and Hepatitis C patients and demographic values like age was used to train a kNN model. For the training of this set the CRISP-DM model was used, as can be seen below. More information about the dataset can be found here:
[KNN-occupancy.csv](https://archive.ics.uci.edu/ml/datasets/Occupancy+Detection+) 

## Business Understanding
text and code here

## Data Understanding
The data used comes from UCI Machine Learning Repository and can be found online as an open source dataset ([UCI Machine Learning Repository: HCV Data Set](https://archive.ics.uci.edu/ml/datasets/Occupancy+Detection+)). This dataset was created by R. Lichtinghagen, F. Klawonn & G. Hoffmann. The dataset contains laboratory values of blood donors and Hepatitis C patients and demographic values like age. The dataset has 14 variables (columns) and 569 observations (rows).  


```{r}
str(RawData)
```

## Data Preparation
To create a clean training set a new dataset was created.

```{r}
CleanDF<-RawData[-1]
head(CleanDF)
```

Next, the non-relevant variables, such as the variable 'X', were deleted from the newly created dataset. 

The variable named 'Category' contains the outcome to predict - '0' for 'Blood Donor', '0s' for 'Suspect Blood Donor' and '1','2' & '3' for 'Hepatitis C patients'.

```{r}
cntDiag <- table(CleanDF$Category)
propDiag <- round(prop.table(cntDiag) * 100 , digits = 1)

cntDiag
```
As can be seen, there are five categories while we only want to predict if someone is a blood donor or a Hepatitis C patient. To clean this column or attribute in the training set we can transform the data in this column:

```{r}
CleanDF$Category <- factor(CleanDF$Category, levels = c("0=Blood Donor","0s=suspect Blood Donor", "1=Hepatitis", "2=Fibrosis", "3=Cirrhosis"), labels = c("Blood Donor", "Blood Donor", "Hepatitis C", "Hepatitis C", "Hepatitis C"))
head(CleanDF, 10)
```
Now lets also look at the missing values, to check if any column has missing values we can use the following command:

```{r}
CleanDF %>% map(~sum(is.na(.)))
```
After running this command it shows that the columns ALB, ALP, ALT, CHOL and PROT have missing values or na (not avalaible) as a value. Lets delete the rows with the missing values to create a clean dataset. 

```{r}
CleanDF_omit <- drop_na(CleanDF)
CleanDF_omit
```

Next up the values in the columns have to be normalized


```{r}

normalize <- function(x) { 
  return((x - min(x)) / (max(x) - min(x))) 
}

CleanDF_norm<-as.data.frame(lapply(CleanDF_omit[4:13], normalize))

CleanDF_3<-subset(CleanDF_omit,select=-c(2:13))

CleanDF_n<-cbind(CleanDF_3,CleanDF_norm)

head(CleanDF_n)

```
```{r}
set.seed(38)

rows<-sample(nrow(CleanDF_n))

RandomDF<-CleanDF_n[rows, ]

head(RandomDF)
```


```{r}
trainDF_feat <- RandomDF[1:471, -1]
testDF_feat <- RandomDF[472:589, -1]

trainDF_labels <- as.data.frame(RandomDF[1:471, 1])
testDF_labels <- as.data.frame(RandomDF[472:589, 1])
```

```{r}
cleanDF_test_pred <- knn(train = as.matrix(trainDF_feat), test = as.matrix(testDF_feat), cl = as.matrix(trainDF_labels), k = 21)
head(cleanDF_test_pred)
```
```{r}
confusionMatrix(cleanDF_test_pred, testDF_labels[[1]], positive = NULL, dnn = c("Prediction", "True"))
```